Preliminary Thesis Defence

### 2014-01-06 uncertainty.tex
include formal specification SA, clear chapter pages
spelling

\begin{refsection}
\chapter{Risk and Uncertainty in Unique Equilibria in Currency Attacks}
\label{unc}

\section*{Abstract}
\textcite{morris1998unique} finds that the introduction of a risk factor leads to a unique equilibrium in their model of self-fulfilling currency attacks. The currency attack model thus has the surprising feature, with the introduction of perception distortion, the multiple equilibria region seems to solve to a single equilibrium. It is our contention that this is not a surpising result, it a consequence of the second-order effect, where the risk factor replaces a previous existent, Knightian uncertainty. Previously, speculators had no expectations on other speculators expectations. Here, the risk factor replaces uncertainty, this is the main driver of the surprising result of the paper. Also, the new quantified-risk region only solves a certain part of the multiple equilibria region, not the whole of it. Lastly, if we replace the somewhat strange assumption of a bounded uniform distribution of risk perceptions, with the more orthodox, normal distribution, we cannot derive the same result.

@@

In \textcite{morris1998unique} the authors describe such a model.
This model is characterised by a situation where there is a stable, unstable,
and a `ripe for attack' region (based on the underlying economic fundamentals).
The authors expand on the standard model by introducting second-order expectations.
Hereby speculators do not only look at the economics fundamentals of a currency,
but also at other speculators' perceptions of these fundamentals.

The authors proceed by introducing a measure of risk around perceptions of the fundamentals by investors.
Investors do not know the actual state of the fundamentals, but rather a distored form of it.
As a result, the investors' perception of other investors' perceptions are even more distorted.
Paradoxically, the leads to model being solvable to a unique equilibrium.

\section{The Model}
\label{unc:model}
I will give a brief description of the model as it is defined, and some it features.
There is a state of economic fundamentals $\theta$ which is distributed as $\theta \sim U[0,1]$.
The exchange rate is the hypothetical situation of no government intervention is a function only of these fundamental ($f(\theta)$),
it is assumed that $\frac{\partial f}{\partial \theta} > 0$.
The currency is pegged at a level larger than that derived from the fundamentals ($e* \geq f(\theta)$).
Speculators can short the currency at a cost $t$, their payoff is described by:
\begin{equation}\label{unc:specPO}
e^* - f(\theta) - t
\end{equation}
The government derives a positive value $\nu > 0$ from defending the peg.
The cost of the peg is determined by the economic fundamentals and the proportion of the speculators ($\alpha$) that attack the currency.
The payoff the government is thus:
\begin{equation}\label{unc:govPO}
\nu - c(\alpha, \theta)
\end{equation}
Lastly, the following, assumptions are imposed: $c(0,0) > \nu$ (with worst fundamtentals the peg always has to be released),
$c(1,1) > \nu$ (with all speculators attacking, the peg always has to be released),
$e^* - f(1) < t$ (with best fundamentals, it is inopportune for speculators to attack).

@@

Using the turning points described above, we can define three possible outcome intervals:
\begin{enumerate}
	\item $[0, \underaccent{\bar}{\theta}]$, the cost is always too high, this is the unstable region.
	\item $[\underaccent{\bar}{\theta}, \bar{\theta}]$, if enough speculators attack, the cost of defending the peg becomes too high,
	this is the `ripe for attack' region.
	\item $[\bar{\theta}, 1]$, the cost of shorting the currency will always outway the possible gains,
	this is the stable region.
\end{enumerate}
The two corner intervals have unique equilibria, however, the middle interval, does not. This is thus the multiple equilibria region.

\section{The Introduction of Risk}
The authors expand the model by introducing a measure of risk for fundamentals:
\begin{equation}
x \sim U[\theta - \epsilon, \theta + \epsilon]
\end{equation}
Where is the $x$ is the state of the fundamentas \textit{observed} by the speculators.
A speculators observation can thus \textit{at most} deviate from the true value of the fundamentals by $\epsilon$.
The speculators are also aware of the nature of the distortion.
This means that they are also aware of the fact that other speculators' perceived value of the fundatmentals can at most,
deviate from their own with a magnitude of $2\epsilon$.

\subsection{The Outcomes with Risk}
\label{unc:results}
Having established the new model, the authors state:
\begin{quotation}
	The unique optimal strategy for the government is then to abandon the exchange rate only if the observed fraction of deviators, $\alpha$, is greater than or equal to the critical mass $a(\theta)$ in the prevailing state $\theta$.\hspace{\stretch{1}}\parencite[p.~591]{morris1998unique}
\end{quotation}
By solving out this government `optimal' strategy,
the speculators can ensure themselves of a certain abandonment of the peg,
for a sufficiently low observed $x$.
The authors proceed to derivde from this the main result of the paper, which is:
\begin{quotation}
	THEOREM 1: The is a unique $\theta^*$ such that,
	in any equilibrium of the game with imperfect information,
	the government abandons the currency peg if and only if $\theta \leq \theta^*$. \hspace{\stretch{1}}\parencite[p.~592]{morris1998unique}
\end{quotation}
We will now discuss the issues we find in the derivation of this result.

@@

\subsection{The `Ripe for Attack' Region}
We can summarise the first problem with the model as follows:
\begin{enumerate}
	\item Multiple equilibria exist in the interval $(\underaccent{\bar}{\theta},\bar{\theta})$.
	\item The introduction of quantief risk, removes uncertainty with $2 \epsilon$ of $x$.
	\item Here $u(x,\theta) > 0$ leads to $\pi (x) = 1$.
	\item Authors claim that i.f.f. $x<k$ (the quantified-risk region) speculators will attack. 
	\item This is assumption does not follow from the model.
	\item Outside the quantified-risk area the previously existing uncertainty remains.
	\item Multiple equilibria remain possible here.
\end{enumerate}
Consider the statement of the authors:
\begin{quotation}
	For the next step,
	consider the strategy profile where every speculator attacks the currency if and only if
	the message $x$ is less than some fixed number $k$.
	\hspace{\stretch{1}}\parencite[p.~592]{morris1998unique}
\end{quotation}
The risk profile of the speculator is being redefined here.
In the first iteration of the model,
speculators might decide to the attack the currency without knowing what other investors would do.
In this second iteration,
speculators will only attack if they believe that there is a sufficient number of other speculators doing the same.
Yet in the first model, attacks were possible, in absense of this certainty.
It thus follows that speculators should be willing to attack the currency,
even in outside the quantified-risk region (which now provides certainty).

To illustrate this, consider the following. The risk factor has a magnitude $\epsilon$,
where it is assumed that $\epsilon > 0$.
However, if in stead we state that $\epsilon = 0$, we have eliminated the risk factor.
By eliminating the risk factor, we return to a state of perfect information.
Therefore, solving the second iteration of the model, with $\epsilon = 0$ should give us the same result as the first iteration.
In stead, we find that we end again in a situation of unique equilibria.

As a consequence of this we have the previously mentioned:
\begin{quotation}
	The unique optimal strategy for the government is then to abandon the exchange rate
	only if the observed fraction of deviators, $\alpha$,
	is greater than or equal to the critical mass $a(\theta)$
	in the prevailing state $\theta$.
	\hspace{\stretch{1}}\parencite[p.~591]{morris1998unique}
\end{quotation}
This should say ``...abandon the exchange rate \textit{always} if the observed fraction is great than or equal to...''.
Since quantified-risk region now gives the speculators certainty about the other speculators behavious,
this will now always lead to an sustainable attack.

However, as described above, outside of the new quantied-risk region, the old `ripe for attack region' remains.
In this region of multiple equilibria, currency attacks may be successful, or they may not be.

The condition the governments action is thus a sufficient, but not a necesarry one.

\subsection{Risk and Uncertainty}
The second problem is the most fundamental. The distortion measure is presented as a function that adds uncertainty.
If we differenciate between Knightian uncertainty and risk, then this is not true.

@@

\subsection{A Bounded Uniform Distortion}
The third problem is relatively straightforward.
The assumption of a uniform distortion is unorthodox.
We would normally expect that these distortion would take the shape of e.g.~a normal distribution.
The effect of this is extremely relevant.
A uniform distribution is bounded,
which allows the distortion to the quantied with absolute certainty.
An unbounded distribution, such as the normal distribution,
would not allow a speculator to know the limit of other speculators perception,
with absolute certainty.

@@

Firstly, the new equilibrium, does not cover the entire previously existing multiple equilibria region.
Even though a larger part of the region (w.r.t.~the economic fundamentals) leads to a unique equilibrium,
a part of the old 'ripe for attack' region, remains subject to a situation of multiple equilibria.
It is only through two assumptions which do not follow from the model,
that this region also solves for a unique equilibrium.

@@

Thirdly, the authors introduce a distortion which takes the form of a uniform distribution.
It is only due to the bounded nature of the uniform distribution,
that speculators are able to form bounded quantitative exectatations on other speculators expectations, with absolute certainty.
As mentioned above, it is this absolute certainty on the bounded quantifiable second-order expectations,
that drives the result.

In conclusion, we analyse the model introducted by \textcite{morris1998unique} and find that the results are driven by a second-order effect, as well as certain unjustifiable assumptions.

>>>>

\begin{refsection}
\chapter{Risk and Uncertainty in Unique Equilibria in Currency Attacks}
\label{unc}
\vspace{\stretch{1}}
\section*{Abstract}
\textcite{morris1998unique} finds that the introduction of a risk factor leads to a unique equilibrium in their model of self-fulfilling currency attacks.
The currency attack model thus has the surprising feature, with the introduction of perception distortion,
the multiple equilibria region seems to solve to a unique equilibrium.

It is our contention that this is not a surprising result, it a consequence of the second-order effect,
where the risk factor replaces a previous existent, Knightian uncertainty.
Previously, speculators had no expectations on other speculators expectations.
Here, the risk factor replaces uncertainty, this is the main driver of the surprising result of the paper.
Also, the new quantified-risk region only solves a certain part of the multiple equilibria region, not the whole of it.
Lastly, if we replace the somewhat strange assumption of a bounded uniform distribution of risk perceptions,
with the more orthodox, normal distribution, we cannot derive the same result.
\pagebreak

@@

In \textcite{morris1998unique} the authors describe such a model.
This model is characterised by a situation where there is a stable, unstable,
and a `ripe for attack' region (based on the underlying economic fundamentals).
The authors expand on the standard model by introducing second-order expectations.
Hereby speculators do not only look at the economics fundamentals of a currency,
but also at other speculators' perceptions of these fundamentals.

The authors proceed by introducing a measure of risk around perceptions of the fundamentals by investors.
Investors do not know the actual state of the fundamentals, but rather a distorted form of it.
As a result, the investors' perception of other investors' perceptions are even more distorted.
Paradoxically, the leads to model being solvable to a unique equilibrium.

\section{The Model}
\label{unc:model}
I will give a brief description of the model as it is defined, and some it features.
There is a state of economic fundamentals $\theta$ which is distributed as $\theta \sim U[0,1]$.
The exchange rate is the hypothetical situation of no government intervention is a function only of these fundamental ($f(\theta)$),
it is assumed that $\frac{\partial f}{\partial \theta} > 0$.
The currency is pegged at a level larger than that derived from the fundamentals ($e* \geq f(\theta)$).
Speculators can short the currency at a cost $t$, their payoff is described by:
\begin{equation}\label{unc:specPO}
e^* - f(\theta) - t
\end{equation}
The government derives a positive value $\nu > 0$ from defending the peg.
The cost of the peg is determined by the economic fundamentals and the proportion of the speculators ($\alpha$) that attack the currency.
The payoff the government is thus:
\begin{equation}\label{unc:govPO}
\nu - c(\alpha, \theta)
\end{equation}
Lastly, the following, assumptions are imposed: $c(0,0) > \nu$ (with worst fundamentals the peg always has to be released),
$c(1,1) > \nu$ (with all speculators attacking, the peg always has to be released),
$e^* - f(1) < t$ (with best fundamentals, it is inopportune for speculators to attack).

@@

Using the turning points described above, we can define three possible outcome intervals:
\begin{enumerate}
	\item $[0, \underaccent{\bar}{\theta}]$, the cost is always too high, this is the unstable region.
	\item $[\underaccent{\bar}{\theta}, \bar{\theta}]$, if enough speculators attack, the cost of defending the peg becomes too high,
	this is the `ripe for attack' region.
	\item $[\bar{\theta}, 1]$, the cost of shorting the currency will always outweigh the possible gains,
	this is the stable region.
\end{enumerate}
The two corner intervals have unique equilibria, however, the middle interval, does not. This is thus the multiple equilibria region.

\section{The Introduction of Risk}
The authors expand the model by introducing a measure of risk for fundamentals:
\begin{equation}
x \sim U[\theta - \epsilon, \theta + \epsilon]
\end{equation}
Where is the $x$ is the state of the fundamentals \textit{observed} by the speculators.
A speculators observation can thus \textit{at most} deviate from the true value of the fundamentals by $\epsilon$.
The speculators are also aware of the nature of the distortion.
This means that they are also aware of the fact that other speculators' perceived value of the fundamentals can at most,
deviate from their own with a magnitude of $2\epsilon$.

\subsection{The Outcomes with Risk}
\label{unc:results}
Having established the new model, the authors state:
\begin{quotation}
	The unique optimal strategy for the government is then to abandon the exchange rate only if the observed fraction of deviators, $\alpha$, is greater than or equal to the critical mass $a(\theta)$ in the prevailing state $\theta$.\hspace{\stretch{1}}\parencite[p.~591]{morris1998unique}
\end{quotation}
By solving out this government `optimal' strategy,
the speculators can ensure themselves of a certain abandonment of the peg,
for a sufficiently low observed $x$.
The authors proceed to derive from this the main result of the paper, which is:
\begin{quotation}
	THEOREM 1: The is a unique $\theta^*$ such that,
	in any equilibrium of the game with imperfect information,
	the government abandons the currency peg if and only if $\theta \leq \theta^*$. \hspace{\stretch{1}}\parencite[p.~592]{morris1998unique}
\end{quotation}
We will now discuss the issues we find in the derivation of this result.

@@

\subsection{The `Ripe for Attack' Region}
We can summarise the first problem with the model as follows:
\begin{enumerate}
	\item Multiple equilibria exist in the interval $(\underaccent{\bar}{\theta},\bar{\theta})$.
	\item The introduction of quantied risk, removes uncertainty with $2 \epsilon$ of $x$.
	\item Here $u(x,\theta) > 0$ leads to $\pi (x) = 1$.
	\item Authors claim that i.f.f. $x<k$ (the quantified-risk region) speculators will attack. 
	\item This is assumption does not follow from the model.
	\item Outside the quantified-risk area the previously existing uncertainty remains.
	\item Multiple equilibria remain possible here.
\end{enumerate}
Consider the statement of the authors:
\begin{quotation}
	For the next step,
	consider the strategy profile where every speculator attacks the currency if and only if
	the message $x$ is less than some fixed number $k$.
	\hspace{\stretch{1}}\parencite[p.~592]{morris1998unique}
\end{quotation}
The risk profile of the speculator is being redefined here.
In the first iteration of the model,
speculators might decide to the attack the currency without knowing what other investors would do.
In this second iteration,
speculators will only attack if they believe that there is a sufficient number of other speculators doing the same.
Yet in the first model, attacks were possible, in absence of this certainty.
It thus follows that speculators should be willing to attack the currency,
even in outside the quantified-risk region (which now provides certainty).

To illustrate this, consider the following. The risk factor has a magnitude $\epsilon$,
where it is assumed that $\epsilon > 0$.
However, if in stead we state that $\epsilon = 0$, we have eliminated the risk factor.
By eliminating the risk factor, we return to a state of perfect information.
Therefore, solving the second iteration of the model, with $\epsilon = 0$ should give us the same result as the first iteration.
In stead, we find that we end again in a situation of unique equilibria.
This thus indicates that the second order effect are not really taken into account in the first iteration of the model.

As a consequence of this we have the previously mentioned:
\begin{quotation}
	The unique optimal strategy for the government is then to abandon the exchange rate
	only if the observed fraction of deviators, $\alpha$,
	is greater than or equal to the critical mass $a(\theta)$
	in the prevailing state $\theta$.
	\hspace{\stretch{1}}\parencite[p.~591]{morris1998unique}
\end{quotation}
This should say ``...abandon the exchange rate \textit{always} if the observed fraction is great than or equal to...''.
Since quantified-risk region now gives the speculators certainty about the other speculators behaviour,
this will now always lead to an sustainable attack.

However, as described above, outside of the new quantified-risk region, the old `ripe for attack region' remains.
In this region of multiple equilibria, currency attacks may be successful, or they may not be.

The condition the governments action is thus a sufficient, but not a necessary one.

\subsection{Risk and Uncertainty}
The second problem is the most fundamental. The distortion measure is presented as a function that adds uncertainty.
If we differentiate between Knightian uncertainty and risk, then this is not true.

@@

\subsection{A Bounded Uniform Distortion}
The third problem is relatively straightforward.
The assumption of a uniform distortion is unorthodox.
We would normally expect that these distortion would take the shape of e.g.~a normal distribution.
The effect of this is extremely relevant.
A uniform distribution is bounded,
which allows the distortion to the quantified with absolute certainty.
An unbounded distribution, such as the normal distribution,
would not allow a speculator to know the limit of other speculators perception,
with absolute certainty.

@@

Firstly, the new equilibrium, does not cover the entire previously existing multiple equilibria region.
Even though a larger part of the region (w.r.t.~the economic fundamentals) leads to a unique equilibrium,
a part of the old `ripe for attack' region, remains subject to a situation of multiple equilibria.
It is only through two assumptions which do not follow from the model,
that this region also solves for a unique equilibrium.

@@

Thirdly, the authors introduce a distortion which takes the form of a uniform distribution.
It is only due to the bounded nature of the uniform distribution,
that speculators are able to form bounded quantitative expectations on other speculators expectations, with absolute certainty.
As mentioned above, it is this absolute certainty on the bounded quantifiable second-order expectations,
that drives the result.

In conclusion, we analyse the model introduced by \textcite{morris1998unique} and find that the results are driven by a second-order effect, as well as certain unjustifiable assumptions.
